\documentclass[main]{subfiles}
\begin{document}
\chapter*{Preface to the new typesetting}
Credits for typesetting go to (in no particular order): Aareyan Manzoor, George Coote, John Cerkan, DarQ, Prakhar Agarwal, Andres Klene, Bastián Núñez, Slurp, Bryan Li, Yohan Wittgenstein, Manan Jain, Carl Sun. %add your names in here

Here is a link to a dyslexic friendly version: \url{https://aareyanmanzoor.github.io/assets/blue-book-dyslexic.pdf}

We changed Adams notation $SX$ for the (reduced) suspension of $X$ to $\Sigma X$, which is more modern. We also changed $\bQ_p$ to $\bZ_{(p)}$ to mean the localization of $\bZ$ at $p$. $f|A$ was used originally to talk about restrictions of functions, however we went with the more modern $f|_A$.

A lot of small typos in the original were fixed here. In addition for bigger ones, a footnote was added to indicate it. The original book did not have any footnotes, so all footnotes are an addition of the new retyping. 

A part \ref{part:p1} bibliography was added, as a lot of citations were made inside the text. Some of the in-text references were added to the bibliography for all parts. All citations now have an URL whenever applicable, to the desired article/book etc. Citations in text are color red.

We added hyperlinks for whenever the book referenced a e.g. theorem of itself. These are always color blue. A table of contents was also added with links.

We added an index with things that Adams highlighted in the original, alongside all the different spectra he mentions, with a link to the appearance in the actual book. This can be extended to be more comprehensive in future revisions, so e-mails with index suggestions will be appreciated.

We also apologize that our cover is not Blue.


\end{document}